---
title: "COVID-19 mortality and vaccination distribution in Austin, Texas"
author: "UT-Austin COVID-19 Modeling Consortium"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

<style type="text/css">
.main-container {
  max-width: 2400px;
  margin-left: auto;
  margin-right: auto;
}
</style>



```{r include=FALSE,fig.width=6, fig.height=6, fig.align='center',message=F}
library(ggplot2)
library(ggplot2)
library(dplyr)
library(cowplot)
library(sf)
library(ggiraph)
## Diverging color scales
mycols1 <- c("#8c510a", "#01665e", "#f5f5f5")
mycols2 <- c("#c51b7d", "#4d9221", "#f7f7f7")
mycols3 <- c("#b2182b", "#2166ac", "#f7f7f7")
mycols4 <- c("#d73027", "#4575b4", "#ffffbf")
mycols5 <- c("#b2182b", "#4d4d4d", "#ffffff")
mycols6 <- c("#b35806", "#542788", "#f7f7f7")
```



```{r include=FALSE, echo=FALSE}
theme_set(theme_half_open())

load("./austin.Rdata")

# ZIP code: 
# Vaccine coverage: XX.X%
# Infection rate: 
# SVI: 
# Mortality rate: XX per 10,000

vax_sub_deaths <- vax_sub_deaths %>% 
  mutate(ZCTA_label = 
           sprintf("ZIP code: %s\nInfection rate: %2.1f%%\nMortality rate: %2.1f per 10,000 residents\nVaccine coverage: %2.1f%%\nSVI: %1.2f", 
                   ZCTA, 
                   round(attack_rate_mean * 100, 1),
                   round(deaths_per_10k, 2),
                   round(coverage * 100, 1),
                   round(SVI, 2)))

vax_sub_deaths <- vax_sub_deaths %>% 
  mutate(ZCTA_label = ZCTA
           # sprintf("ZIP code: %s\nSVI: %1.2f", 
                   # ZCTA, 
                   # round(attack_rate_mean * 100, 1),
                   # round(deaths_per_10k, 2),
                   # round(coverage * 100, 1),
                   # round(SVI, 2))
)

city_avg_deaths <- vax_sub_deaths %>%
  mutate(infected = B01001_001E * attack_rate_mean) %>%
  summarize(
    SVI = mean(SVI),
    total_pop = sum(B01001_001E),
    total_infected = sum(infected, na.rm=TRUE),
    infection_rate = total_infected / total_pop,
    adult_pop = sum(adult_pop),
    one_dose = sum(one_dose),
    senior_frac = sum(plus65)/adult_pop,
    nonsenior_frac=sum(adults_16_64)/adult_pop,
    total_deaths = sum(deaths),
    total_deaths_65plus = sum(deaths_65plus),
    total_deaths_65under = sum(deaths_65under),
    city_pop = sum(total_pop),
    city_pop_65plus = sum(adults_65_plus),
    city_pop_65under = city_pop - city_pop_65plus,
    deaths_per_10k_avg = total_deaths/city_pop * 1e4,
    deaths_per_10k_65plus_avg = total_deaths_65plus / city_pop_65plus * 1e4,
    deaths_per_10k_65under_avg = total_deaths_65under / city_pop_65under * 1e4
  ) %>%
    mutate(coverage = one_dose / adult_pop,
       coverage_all = one_dose / total_pop)

## All deaths
myplot3_deaths_ia <- vax_sub_deaths %>%
  mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>%
  # glimpse() %>%
  ggplot() +
  geom_sf_interactive(
    aes(fill=deaths_per_10k,
        tooltip = ZCTA_label,
        data_id = ZCTA_label,
        label=ZCTA_label,
        geometry=geometry## , col=burden
        ), size=0.1) +
  ## Roads
  geom_sf(data = I35_2,
          col = "grey10") +
  geom_sf(data = missing_183_2,
          col = "grey10") +
  geom_sf(data = US183_2,
          col = "grey10") +
  annotate(geom="text", x=-97.8, y=30.55, label="US 183", size=4)+
  annotate(geom="text", x=-97.68, y=30.55, label="I-35", size=4) +
  scale_fill_viridis_c("deaths per\n10,000 residents",option="D"## labels=scales::label_percent()
  ) +
  NULL +
  labs(title="Mortality rate\nAll ages")  +
  guides(fill = guide_coloursteps(ticks=TRUE,barwidth = 14, barheight = 0.5)) +
  theme_map() +
  theme(legend.position = "top",
        legend.title = element_text(face = "bold"))



mycols_death <- mycols5
mycols_death <- c("#762a83", "#1b7837", "#f7f7f7")

myplot3_deaths_ia2 <- vax_sub_deaths %>%
  mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>%
  # glimpse() %>%
  ggplot() +
  geom_sf_interactive(
    aes(fill=deaths_per_10k,
        tooltip = ZCTA_label,
        data_id = ZCTA_label,
        label=ZCTA,
        geometry=geometry## , col=burden
        ), size=0.1) +
  ## Roads
  geom_sf(data = I35_2,
          col = "grey10") +
  geom_sf(data = missing_183_2,
          col = "grey10") +
  geom_sf(data = US183_2,
          col = "grey10") +
  annotate(geom="text", x=-97.8, y=30.55, label="US 183", size=4)+
  annotate(geom="text", x=-97.68, y=30.55, label="I-35", size=4) +
  scale_fill_steps2(    "",
                    midpoint=round(city_avg_deaths$deaths_per_10k_avg/1)*1,
                    ## , mid="grey90"
                    # high = mycols_death[2], 
                    # low = mycols_death[1], 
                    high = mycols_death[1], 
                    low = mycols_death[2], 
                    # mid="lightgrey",
                    mid = mycols_death[3]
) + 
  # scale_fill_viridis_c(
  #   "",
  #   # "deaths per\n10,000 residents",
  #                      option="D"## labels=scales::label_percent()
  # ) +
  NULL +
  labs(subtitle="Mortality rate per 10,000 residents")  +
  guides(fill = guide_coloursteps(ticks=TRUE,barwidth = 14, barheight = 0.5)) +
  theme_map() +
  theme(legend.position = "top",
        legend.title = element_text(face = "bold"))


# 
# myplot3_deaths_ia2 <- vax_sub_deaths %>%
#   mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>%
#   # glimpse() %>%
#   ggplot() +
#   geom_sf_interactive(
#     aes(fill=deaths_per_10k,
#         tooltip = ZCTA_label,
#         data_id = ZCTA_label,
#         label=ZCTA,
#         geometry=geometry## , col=burden
#         ), size=0.1) +
#   ## Roads
#   geom_sf(data = I35_2,
#           col = "grey10") +
#   geom_sf(data = missing_183_2,
#           col = "grey10") +
#   geom_sf(data = US183_2,
#           col = "grey10") +
#   annotate(geom="text", x=-97.8, y=30.55, label="US 183", size=4)+
#   annotate(geom="text", x=-97.68, y=30.55, label="I-35", size=4) +
#   scale_fill_viridis_c(
#     "",
#     # "deaths per\n10,000 residents",
#                        option="D"## labels=scales::label_percent()
#   ) +
#   NULL +
#   labs(subtitle="Mortality rate per 10,000 residents")  +
#   guides(fill = guide_coloursteps(ticks=TRUE,barwidth = 14, barheight = 0.5)) +
#   theme_map() +
#   theme(legend.position = "top",
#         legend.title = element_text(face = "bold"))

# myplot3_deaths_ia

# save(vax_sub_deaths, file="austin.Rdata")

myscatter41_ia <- vax_sub_deaths %>%
  ggplot() +
  geom_smooth(aes(SVI, deaths_per_10k), method="lm") +
  geom_point_interactive(
    aes(SVI, deaths_per_10k, 
        # label=ZCTA,
        tooltip = ZCTA_label,
        data_id = ZCTA_label,
        color=(total_pop)
        ), size=5,alpha=0.8) +
  scale_color_viridis_c("ZIP population",
                        option="A",direction=-1,
                        breaks = scales::breaks_log(6),
                        labels=scales::label_comma()## ,
                        ## labels = scales::label_percent(accuracy=1)
                        ) +
  labs(subtitle="All ages",
       x = "SVI (higher is more vulnerable)", y = "Deaths per 10,000 residents") +
  theme_grey(base_size=16) +
  guides(color = guide_colorsteps(ticks=TRUE,barwidth = 15, barheight = 0.5)) +
  theme(legend.position = "bottom",
        axis.title = element_text(face="bold"),
        legend.title = element_text(face="bold"))

# myscatter41_ia

```


```{r include=FALSE,echo=FALSE}
myplot3 <- vax_sub_deaths %>%
  mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>% 
  # glimpse() %>% 
  ggplot() +
  geom_sf_interactive(
    aes(fill=coverage, label=ZCTA, geometry=geometry,
        tooltip = ZCTA_label,
        data_id = ZCTA_label## , col=burden
              ), size=0.1) +
  ## Roads
  geom_sf(data = I35_2,
          col = "grey10") +
  geom_sf(data = missing_183_2,
          col = "grey10") +
  geom_sf(data = US183_2,
          col = "grey10") +
  annotate(geom="text", x=-97.8, y=30.55, label="US 183", size=4)+
  annotate(geom="text", x=-97.68, y=30.55, label="I-35", size=4) + 
  scale_fill_steps2(    "",
                    midpoint=round(city_avg_deaths$coverage/0.1)*0.1,## , mid="grey90"
                    high = mycols3[2], low = mycols3[1], mid = mycols3[3],
                    breaks=seq(0, 0.9, by=0.1),
                    labels=scales::label_percent(accuracy=1),
) + 
  NULL +
  labs(subtitle="Vaccine coverage")  +
  guides(fill = guide_coloursteps(ticks=TRUE,barwidth = 15, barheight = 0.5)) +
theme_map() +
## theme_dark() +
  theme(legend.position = "top",
        legend.title = element_text(face = "bold"))

myplot3_svi_mycols <- vax_sub_deaths %>%
  mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>% 
  # glimpse() %>% 
  ggplot() +
  geom_sf_interactive(aes(fill=SVI, label=ZCTA, geometry=geometry,
              tooltip = ZCTA_label,
        data_id = ZCTA_label
              ## , col=burden
              ), size=0.1) +
  ## Roads
  geom_sf(data = I35_2,
          col = "grey10") +
  geom_sf(data = missing_183_2,
          col = "grey10") +
  geom_sf(data = US183_2,
          col = "grey10") +
      scale_fill_steps2(    "",
                        midpoint=round(city_avg_deaths$SVI/0.05)*0.05,## , mid="grey90"
                        high = mycols1[1], low = mycols1[2], mid = mycols1[3],
                    breaks=seq(0, 0.9, by=0.1)## ,
                    ## labels=scales::label_percent(accuracy=1),
) + 
  NULL +
  labs(subtitle="Social vulnerability")  +
  guides(fill = guide_coloursteps(ticks=TRUE,barwidth = 15, barheight = 0.5)) +
  theme_map() +
  annotate(geom="text", x=-97.8, y=30.55, label="US 183", size=4)+
  annotate(geom="text", x=-97.68, y=30.55, label="I-35", size=4) + 
  theme(legend.position = "top",
        legend.title = element_text(face = "bold"))

myplot3_burden <- vax_sub_deaths %>%
  mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>% 
  # glimpse() %>% 
  ggplot() +
  geom_sf_interactive(aes(fill=SVI, label=attack_rate_mean, geometry=geometry,
              tooltip = ZCTA_label,
        data_id = ZCTA_label
              ## , col=burden
              ), size=0.1) +
  ## Roads
  geom_sf(data = I35_2,
          col = "grey10") +
  geom_sf(data = missing_183_2,
          col = "grey10") +
  geom_sf(data = US183_2,
          col = "grey10") +
      scale_fill_steps2(    "",
                        midpoint=round(city_avg_deaths$infection_rate/0.1)*0.1,## , mid="grey90"
                        high = mycols2[1], low = mycols2[2], mid = mycols2[3],
                    breaks=seq(0.1, 0.9, by=0.1),
                    labels=scales::label_percent(accuracy=1),
) + 
  NULL +
  labs(subtitle="Cumulative infections")  +
  guides(fill = guide_coloursteps(ticks=TRUE,barwidth = 15, barheight = 0.5)) +
  theme_map() +
  annotate(geom="text", x=-97.8, y=30.55, label="US 183", size=4)+
  annotate(geom="text", x=-97.68, y=30.55, label="I-35", size=4) + 
  theme(legend.position = "top",
        legend.title = element_text(face = "bold"))

myscatter2 <- vax_sub_deaths %>%
  filter(coverage <= 1) %>% 
  ## mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>% 
  ggplot() +
  geom_smooth(aes(attack_rate_mean, coverage),method="lm") +
  geom_point_interactive(aes(attack_rate_mean, coverage,
                            tooltip = ZCTA_label,
        data_id = ZCTA_label, 
                 col = SVI## , col=B01002_001E
                 ## , col = burden
                 ),
             size = 2.5## , alpha=0.5
             ) +
  ## scale_color_distiller(palette = "Spectral") +
  scale_color_viridis_c("Social vulnerability",
                        option="C") +
  scale_y_continuous(labels=scales::percent) +
  scale_x_continuous(labels=scales::percent) + 
  labs(## title = "Cumulative infections, vaccine coverage, & social vulnerability in Austin", 
       y = "Vaccine coverage",
       x = "Cumulative infections") +
  guides(color = guide_colorsteps(ticks=TRUE,barwidth = 12, barheight = 0.5)) +
  theme(legend.position = "bottom")
  ## theme_cowplot() + 
  # theme(legend.position = "bottom",
  #       axis.title = element_text(face="bold"),
  #       legend.title = element_text(face="bold"))

myscatter_final <- vax_sub_deaths %>%
  filter(coverage <= 1) %>% 
  ## mutate(coverage = ifelse(coverage > 1, 1, coverage)) %>% 
  ggplot() +
  geom_smooth(aes(deaths_per_10k, coverage),method="lm") +
  geom_point_interactive(aes(deaths_per_10k, coverage,
                            tooltip = ZCTA_label,
        data_id = ZCTA_label, 
                 col = SVI## , col=B01002_001E
                 ## , col = burden
                 ),
             size = 2.5## , alpha=0.5
             ) +
  ## scale_color_distiller(palette = "Spectral") +
  scale_color_viridis_c("Social vulnerability",
                        option="C") +
  scale_y_continuous(labels=scales::percent) +
  # scale_x_continuous(labels=scales::percent) + 
  labs(## title = "Cumulative infections, vaccine coverage, & social vulnerability in Austin", 
       y = "Vaccine coverage",
       x = "Mortality rate") +
  guides(color = guide_colorsteps(ticks=TRUE,barwidth = 12, barheight = 0.5)) +
  theme(legend.position = "bottom")

mygrid2 <- plot_grid(
myplot3_svi_mycols, myplot3_burden, myplot3,
      NULL , myscatter2 , NULL,
    nrow=2)

# mygrid2


```






```{r, fig.align="center", fig.width=12, fig.height=10, fig.cap="Figure: Here is a really important caption.", echo=FALSE, message=FALSE}

# girafe(
#   ggobj=mygrid2,
#   # ggobj = plot_grid(
#   #   myplot3_svi_mycols, myplot3_burden, myplot3,
#   #   nrow=1),
#   width_svg = 12,
#   height_svg = 10,
#     options = list(
#     opts_hover_inv(css = "opacity:0.8;"),
#     opts_hover(css = "stroke-width:2;")
#   ))
```



```{r fig.width=12, fig.height=6, fig.align='center', echo=FALSE, message=FALSE}

# ZIP code: 
# Vaccine coverage: XX.X%
# Infection rate: 
# SVI: 
# Mortality rate: XX per 10,000

# How to handle >100% coverage?

# girafe(
#   ggobj = plot_grid(myplot3_deaths_ia, myscatter41_ia),
#   width_svg = 12,
#   height_svg = 6,
#     options = 
#     # list(opts_tooltip(use_fill = TRUE),
#     #              opts_sizing(width = .7))
#   list(
#     opts_hover_inv(css = "opacity:0.2;"),
#     opts_hover(css = "stroke:magenta;stroke-width:2;")
#     # opts_tooltip(use_fill = TRUE),
#     #              opts_sizing(width = .7)
#   )
#   )



  ##   options = list(
  ##   opts_hover(css = "fill: #202020; opacity: 0.8"),
  ##   opts_hover_inv(css = "opacity: 0.8;")
  ## )


```

## Mapping COVID-19 outcomes, vaccine administration, and social vulnerability in Austin

```{r fig.width=12, fig.height=11, fig.align='center', echo=FALSE, message=FALSE}

mygrid_all <- plot_grid(
  ## Infection rate
myplot3_burden, 
## Mortality rate
myplot3_deaths_ia2,
## SVI
myplot3_svi_mycols, 
## Vaccine coverage
myplot3,
  nrow=2)

girafe(
  ggobj = mygrid_all,
  width_svg = 12,
  height_svg = 11,
    options = 
  list(
    opts_hover_inv(css = "opacity:0.2;"),
    opts_hover(css = "stroke:black;stroke-width:1.2;")
  )
  )


```

## More vulnerable areas have higher mortality and less vaccinations

```{r fig.width=4, fig.height=4, fig.align='center', echo=FALSE, message=FALSE}


girafe(
  ggobj = myscatter_final,
  width_svg = 7,
  height_svg = 5,
    options = 
  list(
    opts_hover_inv(css = "opacity:0.2;"),
    opts_hover(css = "stroke:black;stroke-width:1.2;")
  )
  )
```

## Credits

Spencer Woody, Emily Javan, Kaitlyn Johnson, Remy Pasco, Maureen Johnson-León, Michael Lachmann, Spencer J. Fox, and Lauren Ancel Meyers

[UT-Austin COVID-19 Modeling Consortium](https://covid-19.tacc.utexas.edu)

## Contact

[Spencer Woody](https://spencerwoody.github.io)  
Email to `spencer.woody [at] utexas [dot] edu`

Download the code and data [here][ghdata].

[ghdata]: https://github.com/spencerwoody/texvax/tree/main/ here
